Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-10 manuscript/manuscript.Rmd
Fix equation

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
extra_dependencies: ["amsmath", "bm"]

---

@@

\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}

where the regression parameters (weights) for each layer $$k \in [1;K]$$ are represented by the matrices $a_k$ and the 
intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear 
regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the tanh function or 
the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a 
loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the 
predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of 
the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation).
Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

When applying embeddings in a data modeling context using neural networks, the values of the embeddings will be 
calibrated during the same fitting process that calibrates the parameters of the neural network.

### Attention {#attention}
Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modelling tasks. Attention mechanisms allow deep neural nets to choose covariates entering the model in a flexible manner; these types of applications are demonstrated in, for example, [@Vaswani2017a]. In this section we provide an introduction to attention mechanisms and then discuss applications of attention to modeling tabular data .

@@

  ```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times q) \times (q \times d)} \to \mathbb{R}^{1 \times d}, ~~
  (\bold{q}, K) ~\mapsto ~ \mathbf{a} = \langle \mathbf{q}, K^T \mathbb{R}angle.
\label{eq:attention_scores}
\end{equation}
```
The attention scores are then normalized using a soft max function i.e.

@@

  ```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times d) \times (d \times q)} \to \mathbb{R}^{1 \times q}, ~~
  (\mathbf{a^*}, V) ~\mapsto ~ \mathbf{v} = \langle \mathbf{a^*}, V \mathbb{R}angle.
\label{eq:attention_output}
\end{equation}
```

# Predictive Modeling with Embeddings

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is 
made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the 
complete dataset contains over two million transactions, for the purposes of our experiments we limit ourselves to data 
from 2000 to August 2019, which amounts to approximately 1.4 million claims. The dataset can be downloaded from 
Cellar[^10] The dataset includes a rich variety of variables, from occupancy type to coarse coordinates. For our models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

[^9]: https://www.fema.gov/about/reports-and-data/openfema
[^10]: https://cellar.kasa.ai

@@

-   Target variable: Amount paid on building claim
-   Predictors:
    - $\log$ `building_insurance_coverage`
    - `basement_enclosure_type`, 
    - `number_of_floors_in_the_insured_building`
    - *prefix* of `flood_zone`
    - `primary_residence`
-   Link function: $\log$
-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to 
match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take 
the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^8]. For example, `"A01"`, 
`"A02"`, and so on are recoded as simply `"A"`.

[^8]: https://www.floodsmart.gov/flood-map-zone/find-yours

>>>>

---
title: Categorical Predictor Embeddings in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
extra_dependencies: ["amsmath", "bm"]
---

@@

```{=tex}
\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}
```
where the regression parameters (weights) for each layer $$k \in [1;K]$$ are represented by the matrices $a_k$ and the intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the tanh function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation). Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

When applying embeddings in a data modeling context using neural networks, the values of the embeddings will be calibrated during the same fitting process that calibrates the parameters of the neural network. The following equation shows how the first layer of a neural network incorporating embeddings might be written:

```{=tex}
\begin{equation}
  z^1 = \sigma(a_1.X'+ b_1), 
\end{equation}
```
where we represent the output of embedding layers concatenated together with any numerical inputs as $X'$.

### Attention {#attention}

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modelling tasks. Attention mechanisms allow deep neural nets to choose covariates entering the model in a flexible manner; these types of applications are demonstrated in, for example, [@Vaswani2017a]. In this section we provide an introduction to attention mechanisms and then discuss applications of attention to modeling tabular data.

@@

```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times q) \times (q \times d)} \to \mathbb{R}^{1 \times d}, ~~
(\bold{q}, K) ~\mapsto ~ \mathbf{a} = \langle \mathbf{q}, K^T \mathbb\rangle.
\label{eq:attention_scores}
\end{equation}
```
The attention scores are then normalized using a soft max function i.e.

@@

```{=tex}
\begin{equation}
\mathbb{R}^{(1 \times d) \times (d \times q)} \to \mathbb{R}^{1 \times q}, ~~
(\mathbf{a^*}, V) ~\mapsto ~ \mathbf{v} = \langle \mathbf{a^*}, V \mathbb\rangle.
\label{eq:attention_output}
\end{equation}
```
# Predictive Modeling with Embeddings

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments we limit ourselves to data from 2000 to August 2019, which amounts to approximately 1.4 million claims. The dataset can be downloaded from Cellar[^2] The dataset includes a rich variety of variables, from occupancy type to coarse coordinates. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

[^1]: <https://www.fema.gov/about/reports-and-data/openfema>

[^2]: <https://cellar.kasa.ai>

@@

-   Target variable: Amount paid on building claim

-   Predictors:

    -   $\log$ `building_insurance_coverage`
    -   `basement_enclosure_type`,
    -   `number_of_floors_in_the_insured_building`
    -   *prefix* of `flood_zone`
    -   `primary_residence`

-   Link function: $\log$

-   Distribution: gamma

We take the log of the continuous predictor `building_insurance_coverage`, which allows the scale of the predictor to match that of the target variable. Because the `flood_zone` variable in the original data contains 60 levels, we take the prefix of the zone code, which corresponds to the level of risk as determined by FEMA [^3]. For example, `"A01"`, `"A02"`, and so on are recoded as simply `"A"`.

[^3]: <https://www.floodsmart.gov/flood-map-zone/find-yours>
